### 2018-07-17 sections/introduction.tex
Implemented additional CDS comments by Michele Webber

The searches presented in this paper are focused on the dominant fermionic decay modes of the Higgs boson.
Therefore, they help complete the programme of searches for $t \to Hq$ decays at the ATLAS experiment based on proton-proton collision 
data at $\sqrt{s}=13~\tev$ recorded in 2015 and 2016. The corresponding integrated luminosity is 36.1 fb$^{-1}$.
Two analyses are performed searching for $\ttbar \to WbHq$ production, and targeting the $H \to b\bar{b}$ and $H \to \tau^+\tau^-$ decay modes, 
which will be referred to as ``$\Hbb$ search" and ``$\Htautau$ search", respectively.
The $\Hbb$ search selects events with one isolated light lepton from the $W \to \ell\nu$ decay, and multiple jets, with several 
of them being identified with high purity as originating from the hadronisation of $b$-quarks. 
The $\Htautau$ search selects events with either one or two hadronically decaying $\tau$-lepton candidates, as well multiple jets. 
Both searches employ different multivariate techniques to discriminate between the signal and the background on the basis of their kinematic differences. 
These two searches are combined with previous ATLAS searches in the diphoton and multilepton final states using the same dataset~\cite{Aaboud:2017mfd,Aaboud:2018pob}, 
and bounds are set on $\BR(t\to Hc)$ and $\BR(t\to Hu)$, as well as on the corresponding non-flavour-diagonal Yukawa couplings.

>>>>

The searches presented in this paper are focused on the dominant fermionic decay modes of the Higgs boson.
Therefore, they help complete the programme of searches for $t \to Hq$ decays at the ATLAS experiment based on proton-proton collision 
data at $\sqrt{s}=13~\tev$ recorded in 2015 and 2016. The corresponding integrated luminosity is 36.1 fb$^{-1}$.
Two analyses are performed searching for $\ttbar \to WbHq$ production, and targeting the $H \to b\bar{b}$ and $H \to \tau^+\tau^-$ decay modes, 
which will be referred to as ``$\Hbb$ search" and ``$\Htautau$ search", respectively.
The $\Hbb$ search selects events with one isolated light lepton from the $W \to \ell\nu$ decay, and multiple jets, with several 
of them being identified with high purity as originating from the hadronisation of $b$-quarks. 
The $\Htautau$ search selects events with either one or two hadronically decaying $\tau$-lepton candidates, as well multiple jets. 
Both searches employ different multivariate techniques to discriminate between the signal and the background on the basis of their different kinematics. 
These two searches are combined with previous ATLAS searches in the diphoton and multilepton final states using the same dataset~\cite{Aaboud:2017mfd,Aaboud:2018pob}, 
and bounds are set on $\BR(t\to Hc)$ and $\BR(t\to Hu)$, as well as on the corresponding non-flavour-diagonal Yukawa couplings.

### 2018-07-17 sections/statistical_analysis.tex
Implemented additional CDS comments by Michele Webber

For each search, the final discriminant distributions across all analysis regions considered are jointly analysed to test for the 
presence of a signal. The statistical analysis uses a binned likelihood function ${\cal L}(\mu,\theta)$ constructed as
a product of Poisson probability terms over all bins considered in the search. This function depends
on the signal-strength parameter $\mu$, defined as a multiplicative factor to the yield for $\Hq$ signal events
normalised to a reference branching ratio $\BR_{\mathrm{ref}}(t\to Hq)=1\%$,
and $\theta$, a set of nuisance parameters that encode the effect of systematic uncertainties in the signal and background expectations. 
Therefore, the expected total number of events in a given bin depends on $\mu$ and $\theta$. With the exception of a few parameters 
that control the normalisation of the some of the background components (e.g. the \ttbin\ background in the case of the $\Hbb$ search) 
and are treated as free parameters in the fit,  all other nuisance parameters are implemented in the likelihood function 
as Gaussian or log-normal constraints.

@@

The test statistic $q_\mu$ is defined as the profile likelihood ratio: 
$q_\mu = -2\ln({\cal L}(\mu,\hat{\hat{\theta}}_\mu)/{\cal L}(\hat{\mu},\hat{\theta}))$,
where $\hat{\mu}$ and $\hat{\theta}$ are the values of the parameters that
maximise the likelihood function (subject to the constraint $0\leq \hat{\mu} \leq \mu$), and $\hat{\hat{\theta}}_\mu$ are the values of the
nuisance parameters that maximise the likelihood function for a given value of $\mu$. 
The test statistic $q_\mu$ is evaluated with the {\textsc RooFit} package~\cite{Verkerke:2003ir,RooFitManual}.
A related statistic is used to determine the probability that the observed data are compatible with the background-only hypothesis (i.e.~the discovery test)  
by setting $\mu=0$ in the profile likelihood ratio and leaving $\hat{\mu}$ unconstrained: $q_0 = -2\ln({\cal L}(0,\hat{\hat{\theta}}_0)/{\cal L}(\hat{\mu},\hat{\theta}))$.
The $p$-value (referred to as $p_0$) representing the probability of the data being compatible with the 
background-only hypothesis is estimated by integrating
the distribution of $q_0$ from background-only pseudo-experiments, approximated using the asymptotic formulae given in Refs.~\cite{Cowan:2010js,ErratumCowan:2010js}, above the observed value of $q_0$. The observed $p_0$-value is checked for each explored signal scenario.
In the absence of any significant excess above the background expectation, upper limits on $\mu$, and thus on 
$\BR(t\to Hq)$, are derived by using $q_\mu$ in the CL$_{\textrm{s}}$ method~\cite{Junk:1999kv,Read:2002hq}.
For a given signal scenario, values of the $\BR(t\to Hq)$ yielding CL$_{\textrm{s}} < 0.05$, 
where CL$_{\textrm{s}}$ is computed using the asymptotic approximation~\cite{Cowan:2010js,ErratumCowan:2010js}, are excluded at $\geq 95\%$ CL.

>>>>

For each search, the final discriminant distributions across all analysis regions considered are jointly analysed to test for the 
presence of a signal. The statistical analysis uses a binned likelihood function ${\cal L}(\mu,\theta)$ constructed as
a product of Poisson probability terms over all bins considered in the search. This function depends
on the signal-strength parameter $\mu$, defined as a multiplicative factor to the yield for $\Hq$ signal events
normalised to a reference branching ratio $\BR_{\mathrm{ref}}(t\to Hq)=1\%$,
and $\theta$, a set of nuisance parameters that encode the effect of systematic uncertainties in the signal and background expectations. 
Therefore, the expected total number of events in a given bin depends on $\mu$ and $\theta$. 
All nuisance parameters are implemented in the likelihood function as Gaussian or log-normal constraints, with the exception of a few parameters 
that control the normalisation of the some of the background components (e.g. the \ttbin\ background in the case of the $\Hbb$ search) 
and are treated as free parameters in the fit.

@@

The test statistic $q_\mu$ is defined as the profile likelihood ratio: 
$q_\mu = -2\ln({\cal L}(\mu,{\hat{\theta}}_\mu)/{\cal L}(\hat{\mu},\hat{\theta}))$,
where $\hat{\mu}$ and $\hat{\theta}$ are the values of the parameters that
maximise the likelihood function (subject to the constraint $0\leq \hat{\mu} \leq \mu$), and ${\hat{\theta}}_\mu$ are the values of the
nuisance parameters that maximise the likelihood function for a given value of $\mu$. 
The test statistic $q_\mu$ is evaluated with the {\textsc RooFit} package~\cite{Verkerke:2003ir,RooFitManual}.
A related statistic is used to determine the probability that the observed data are compatible with the background-only hypothesis (i.e.~the discovery test)  
by setting $\mu=0$ in the profile likelihood ratio and leaving $\hat{\mu}$ unconstrained: $q_0 = -2\ln({\cal L}(0,{\hat{\theta}}_0)/{\cal L}(\hat{\mu},\hat{\theta}))$.
The $p$-value (referred to as $p_0$) representing the probability of the data being compatible with the 
background-only hypothesis is estimated by integrating
the distribution of $q_0$ obtained from background-only pseudo-experiments, approximated using the asymptotic formulae given in Refs.~\cite{Cowan:2010js,ErratumCowan:2010js}, above the observed value of $q_0$ in the data. The observed $p_0$-value is checked for each explored signal scenario.
In the absence of any significant excess above the background expectation, upper limits on $\mu$, and thus on 
$\BR(t\to Hq)$, are derived by using $q_\mu$ in the CL$_{\textrm{s}}$ method~\cite{Junk:1999kv,Read:2002hq}.
For a given signal scenario, values of the $\BR(t\to Hq)$ yielding CL$_{\textrm{s}} < 0.05$, 
where CL$_{\textrm{s}}$ is computed using the asymptotic approximation~\cite{Cowan:2010js,ErratumCowan:2010js}, are excluded at $\geq 95\%$ CL.

### 2018-07-17 sections/strategy_Hbb.tex
Implemented additional CDS comments by Michele Webber

Given the choice of the $W\to\ell\nu$ and $H\to b\bar{b}$ decay modes, the $\ttbar \to WbHq$ signal 
is expected to have four jets and three $b$-quarks in the final state, which can be effectively exploited 
to suppress the background. 
Additional jets can also be present because of initial- or final-state radiation.
However, the use of the 60\% $b$-tagging operating point, characterised by a low mistag rate for
charm and light jets, make both $\Hc$ and $\Hu$ signals have a similar $b$-tag multiplicity spectrum,
with a very small fraction of events having four or more $b$-tagged jets.

@@

After event categorisation, the signal-to-background ratio is very low even in the most sensitive analysis regions, and a suitable
discriminating variable between signal and background needs to be constructed in order to improve the sensitivity of the search.
Since both signal and background result from the $\ttbar$ decay, there are few experimental handles available to discriminate
between them. The most prominent features are the different resonances present in the decay (i.e. the Higgs boson in the case 
of $\Hq$ signal and a hadronically decaying $W$ boson in the case of the $\ttbar \to WbWb$ background), and the different flavour content of the 
jets forming those resonances. However, the large number of jets in the final state causes ambiguities in the calculation 
of these kinematic variable to discriminate signal events from background events.

@@

Two background hypotheses are considered, corresponding to the dominant backgrounds in
the analysis: $\ttbar$+light-jets and $\ttbin$. Thus, $P^\textrm{bkg}$ is computed as the average of
the pdfs for both hypotheses, weighted by their relative fractions in simulated $\ttbar$+jets events, which depend
on the analysis region considered. Furthermore, in a significant fraction of $\Hq$ simulated events (e.g. about 40--50\% in 3b regions), 
the light-quark jet from the hadronic top-quark decay is not among the selected jets.
Similarly, in about 30--40\% (50--90\%) of simulated $\ttbar$+light-jets ($\ttbin$) background events in 3b regions, 
the light-quark jet originating from the $W$ boson decay is also not selected. Thus, the calculation of $P^\textrm{sig}$ and
$P^\textrm{bkg}$ also includes an additional hypothesis to account for this topology, again weighted by the corresponding fractions. 
In this case, the invariant masses involving the missing jet are computed using the highest-$\pt$ jet not matched 
to a decay product from the $\ttbar$ system.

An example of the discrimination between signal and background that is obtained is shown in Figure~\ref{fig:LHD},
where the shape of the LH discriminant distribution is compared between the $\Hc$ and $\Hu$ signals and the 
$t\bar{t}\to WbWb$ background in each of the analysis regions considered.

>>>>

Given the choice of the $W\to\ell\nu$ and $H\to b\bar{b}$ decay modes, the $\ttbar \to WbHq$ signal 
is expected to have four jets in the final state, three of them originating from $b$-quarks, which 
can be effectively exploited to suppress the background. 
Additional jets can also be present because of initial- or final-state radiation.
However, the use of the 60\% $b$-tagging operating point, characterised by a low mistag rate for
charm and light jets, make both $\Hc$ and $\Hu$ signals have a similar $b$-tag multiplicity spectrum,
with a very small fraction of events having four or more $b$-tagged jets.

@@

After event categorisation, the signal-to-background ratio is insufficient even in the best cases to achieve sensitivity, and a suitable
discriminating variable between signal and background needs to be constructed in order to improve the sensitivity of the search.
Since both signal and background result from the $\ttbar$ decay, there are few experimental handles available to discriminate
between them. The most prominent features are the different resonances present in the decay (i.e. the Higgs boson in the case 
of $\Hq$ signal and a hadronically decaying $W$ boson in the case of the $\ttbar \to WbWb$ background), and the different flavour content of the 
jets forming those resonances. However, the large number of jets in the final state causes ambiguities in the calculation 
of these kinematic variable to discriminate signal events from background events.

@@

Two background hypotheses are considered, corresponding to the dominant backgrounds in
the analysis: $\ttbar$+light-jets and $\ttbin$. Thus, $P^\textrm{bkg}$ is computed as the average of
the pdfs for both hypotheses, weighted by their relative fractions found in simulated $\ttbar$+jets events, which depend
on the analysis region considered. Furthermore, in a significant fraction of $\Hq$ simulated events (e.g. about 40--50\% in 3b regions), 
the light-quark jet from the hadronic top-quark decay is not among the selected jets.
Similarly, in about 30--40\% (50--90\%) of simulated $\ttbar$+light-jets ($\ttbin$) background events in 3b regions, 
the light-quark jet originating from the $W$ boson decay is also not selected. Thus, the calculation of $P^\textrm{sig}$ and
$P^\textrm{bkg}$ also includes an additional hypothesis to account for this topology, again weighted by the corresponding fractions. 
In this case, the invariant masses involving the missing jet are computed using the highest-$\pt$ jet not matched 
to a decay product from the $\ttbar$ system.

%An example of the discrimination between signal and background that is obtained is shown in Figure~\ref{fig:LHD},
%where the shape of the LH discriminant distribution is compared between the $\Hc$ and $\Hu$ signals and the 
%$t\bar{t}\to WbWb$ background in each of the analysis regions considered.

Since this analysis has higher expected sensitivity to a $\Hc$ signal than to a $\Hu$ signal and, in order to allow probing 
of the $\BR(t\to Hu)$ versus $\BR(t\to Hc)$ plane, the discriminant optimised for $\Hc$ is used for both 
decay modes. It was verified that using the $\Hc$ discriminant for the $\Hu$ search does not result in a significant sensitivity loss.
Figure~\ref{fig:LHD} compares the shape of the LH distribution between the $\Hc$ and $\Hu$ signals and the 
$t\bar{t}\to WbWb$ background in each of the analysis regions considered.

### 2018-07-17 sections/strategy_Htautau.tex
Implemented additional CDS comments by Michele Webber

In the $\Htautau$ search, the final state signature of interest is characterised by the presence of two $\tau$-leptons from the
decay of the Higgs boson and a light-quark jet, both originating from the decay of one of the top quarks. If one of the $\tau$-leptons
decays leptonically, an isolated electron or muon and significant $\met$ would also be expected.
The other top quark decays into a $W$ boson and a $b$-quark, with the $W$ boson in turn decaying hadronically, giving 
three additional jets, one of which should be tagged as a $b$-jet. Therefore, at least four jets are typically expected in signal events.
However, in a significant fraction of the events the trailing jet from the $W$-boson decay fails the minimum $\pt$ requirement of $30~\gev$,
resulting in signal events with only three jets reconstructed.
In order to optimise the sensitivity of the search, the selected events are categorised into four search 
regions depending on the number of $\taulep$ and $\tauhad$ candidates, and on the number of jets:
($\lephad$, 3j), ($\lephad$, $\geq$4j), ($\hadhad$, 3j), and ($\hadhad$, $\geq$4j).

This event categorisation is primarily motivated by the different quality of the event kinematic reconstruction, depending on the amount 
of $\met$ in the event (larger in $\lephad$ events compared to $\hadhad$ events), and whether a jet from the hadronic top-quark decay 
is missing or not (3j vs $\geq$4j events). This analysis adopts the event reconstruction strategy developed in Ref.~\cite{Chen:2015nta},
which is briefly summarised in the following. Events with exactly three jets that are compatible with having a fully reconstructed hadronically decaying 
top quark ($t \to Wb \to qqb$) necessarily miss the light-quark jet from the $t \to Hq$ decay and are thus rejected. 
For both 3j and $\geq$4j events, the selected jets are assigned to the different top quark decay products via a criterion based on 
minimising a sum of angular distances between objects. Finally, the four-momenta of the invisible decay products for each $\tau$-lepton decay 
are estimated by minimising a $\chi^2$ function based on the probability density functions for the angular distance of the visible and invisible
decay products in the $\tau$-lepton decay, and including Gaussian constraints on the $\tau$-lepton mass, the Higgs boson mass and the
measured $\met$. After the minimisation, an improved resolution is obtained on the determination of the Higgs boson four-momentum, and hence
its invariant mass, as well as on the four-momentum of the parent top quark. Following the event kinematic reconstruction, several kinematic variables
that discriminate between signal and background are defined. These variables are used in the multivariate analysis discussed in the next section.

@@

%%%%%%%%%%%%%%%
\begin{table*}[t!]
\begin{center}
\begin{tabular}{ccccc}
\toprule\toprule 
& \multicolumn{2}{c}{$\lephad$} & \multicolumn{2}{c}{$\hadhad$} \\
Variable & 3j & $\geq$4j & 3j & $\geq$4j  \\
\midrule
$m_{\tau\tau}^{\text{fit}}$                      	& $\times$  & $\times$  & $\times$  & $\times$ \\
$m_{Hq}$                                	& $\times$  & $\times$  & $\times$  & $\times$ \\
$m_{\text{T,lep}}$                              	& $\times$  & $\times$  &             & \\
$p_{\text{T,1}}$                             	&             &             & $\times$  & $\times$ \\
$p_{\text{T,2}}$                             	& $\times$  & $\times$  & $\times$  & $\times$ \\
$\met$ $\phi$ centrality                             	& $\times$  & $\times$  & $\times$  & $\times$ \\
$E_{\text{T},\parallel}^{\text{miss}}$          	& $\times$  & $\times$  & $\times$  & $\times$ \\
$E_{\text{T},\perp}^{\text{miss}}$          	& $\times$  & $\times$  &             & \\
$m_{b j_1}$                   		        & $\times$  & $\times$  & $\times$  & $\times$ \\
$m_{\text{lep}j}$      			        & $\times$  & $\times$  &   	  & \\
$m_{\tau j}$      			        & $\times$  & $\times$  &   	  & \\
$x_1^{\text{fit}}$				& $\times$  &	$\times$  & $\times$  & $\times$ \\
$x_2^{\text{fit}}$				& $\times$  & $\times$  & $\times$  & $\times$ \\
$m_{b j_1 j_2 }$             	   		&             & $\times$  &             & $\times$ \\ \hline
\bottomrule\bottomrule
\end{tabular}
\caption{\small{$\Htautau$ search: Discriminating variables used in the training of the BDT for each search region (denoted by $\times$). 
The description of each variable is provided in the text.}}
\label{tab:mva_var}
\end{center}
\end{table*}
%%%%%%%%%%%%%%%

Boosted decision trees (BDT)~\cite{Breiman:1984jka,Friedman:2002we,Freund:1997xna} are used in each SR to improve the separation between signal and background. The separate training naturally exploits differences in event kinematics across SRs.  
A large set of potential variables was investigated in each SR separately, and only those variables that led to an improved discrimination
performance of the BDT were kept. In the training, only the $\Hq$ signal is used against the total SM background, including both real
and fake $\tauhad$ contributions. The BDT input variables used per SR are listed in Table~\ref{tab:mva_var} and defined in the following:

>>>>

In the $\Htautau$ search, the $\ttbar \to WbHq$ signal being probed is characterised by the presence of $\tau$-leptons from the decay of 
the Higgs boson and at least four jets, only one of which originates from a $b$-quark.
%In the $\Htautau$ search, the final state signature of interest is characterised by the presence of two $\tau$-leptons from the
%decay of the Higgs boson as well as a light-quark jet, the latter two originating from the decay of one of the top quarks. 
If one of the $\tau$-leptons decays leptonically, an isolated electron or muon and significant $\met$ is also expected.
%The other top quark decays into a $W$ boson and a $b$-quark, with the $W$ boson in turn decaying hadronically, resulting in 
%three additional jets. Therefore, at least four jets are typically expected in signal events.
However, in a significant fraction of the events the trailing jet from the $W$-boson decay fails the minimum $\pt$ requirement of $30~\gev$,
resulting in signal events with only three jets reconstructed.
In order to optimise the sensitivity of the search, the selected events are categorised into four SRs depending 
on the number of $\taulep$ and $\tauhad$ candidates, and on the number of jets:
($\lephad$, 3j), ($\lephad$, $\geq$4j), ($\hadhad$, 3j), and ($\hadhad$, $\geq$4j).

This event categorisation is primarily motivated by the different quality of the event kinematic reconstruction, depending on the amount 
of $\met$ in the event (larger in $\lephad$ events compared to $\hadhad$ events), and whether a jet from the hadronic top-quark decay 
is missing or not (3j vs $\geq$4j events). 
%This analysis adopts the event reconstruction strategy developed in Ref.~\cite{Chen:2015nta}, which is briefly summarised in the following. 
The event reconstruction strategy used~\cite{Chen:2015nta} is summarised in the following.
Events with exactly three jets that are compatible with having a fully reconstructed hadronically decaying 
top quark ($t \to Wb \to qqb$) are rejected, as the $t \to Hq$ decay cannot be reconstructed due to the missing light-quark jet.
For both 3j and $\geq$4j events, the selected jets are assigned to the different top quark decay products via a criterion based on 
minimising a sum of angular distances between objects. Finally, the four-momenta of the invisible decay products for each $\tau$-lepton decay 
are estimated by minimising a $\chi^2$ function based on the probability density functions for the angular distance of the visible and invisible
decay products in the $\tau$-lepton decay, and including Gaussian constraints on the $\tau$-lepton mass, the Higgs boson mass and the
measured $\met$. After the minimisation, an improved resolution is obtained on the determination of the Higgs boson four-momentum, and hence
its invariant mass, as well as on the four-momentum of the parent top quark. Following the event kinematic reconstruction, several kinematic variables
that discriminate between signal and background are defined. These variables are used in the multivariate analysis discussed in the next section.

@@

%%%%%%%%%%%%%%%
\begin{table*}[t!]
\begin{center}
\begin{tabular}{ccccc}
\toprule\toprule 
& \multicolumn{2}{c}{$\lephad$} & \multicolumn{2}{c}{$\hadhad$} \\
Variable & 3j & $\geq$4j & 3j & $\geq$4j  \\
\midrule
$m_{\tau\tau}^{\text{fit}}$                      	& $\times$  & $\times$  & $\times$  & $\times$ \\
$m_{Hq}$                                	& $\times$  & $\times$  & $\times$  & $\times$ \\
$m_{\text{T,lep}}$                              	& $\times$  & $\times$  &             & \\
$p_{\text{T,1}}$                             	& $\times$  & $\times$ & $\times$  & $\times$ \\
$p_{\text{T,2}}$                             	& $\times$  & $\times$  & $\times$  & $\times$ \\
$\met$ $\phi$ centrality                             	& $\times$  & $\times$  & $\times$  & $\times$ \\
$E_{\text{T},\parallel}^{\text{miss}}$          	& $\times$  & $\times$  & $\times$  & $\times$ \\
$E_{\text{T},\perp}^{\text{miss}}$          	& $\times$  & $\times$  &             & \\
$m_{b j_1}$                   		        & $\times$  & $\times$  & $\times$  & $\times$ \\
$m_{\text{lep}j}$      			        & $\times$  & $\times$  &   	  & \\
$m_{\tau j}$      			        & $\times$  & $\times$  &   	  & \\
$x_1^{\text{fit}}$				& $\times$  &	$\times$  & $\times$  & $\times$ \\
$x_2^{\text{fit}}$				& $\times$  & $\times$  & $\times$  & $\times$ \\
$m_{b j_1 j_2 }$             	   		&             & $\times$  &             & $\times$ \\ \hline
\bottomrule\bottomrule
\end{tabular}
\caption{\small{$\Htautau$ search: Discriminating variables used in the training of the BDT for each search region (denoted by $\times$). 
The description of each variable is provided in the text.}}
\label{tab:mva_var}
\end{center}
\end{table*}
%%%%%%%%%%%%%%%

Boosted decision trees (BDT)~\cite{Breiman:1984jka,Friedman:2002we,Freund:1997xna} are used in each SR to improve the separation between signal and background. The separate training exploits differences in event kinematics across SRs.  In the training, only the $\Hq$ signal is used against the total SM background, including both real and fake $\tauhad$ contributions.
A large set of potential variables was investigated in each SR separately, and only those variables that led to an improved discrimination
performance of the BDT were kept.  The BDT input variables in each SR are listed in Table~\ref{tab:mva_var} and defined in the following:
